### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@be7007aaeef2c74ad180c22b4d5e984cbccdf2b1 🚀

example on diagnostic test: rare events are hard to detect even with
really good models

summary of model challenges for networks - Strydom et al. (2021)
importance of drawing on traits + validation is challenging - Whalen et
al. (2021) machine learning from genomics

@@

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence. In a previous
work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
of model performance than the True Skill Statistic (TSS), which we will
refer to as Youden's informedness (or \(J\)); these conclusions were
later criticized by Somodi et al. (2017), who emphasized that
informedness' relationship to prevalence depends on assumptions about
bias in the model. Although this work offers recommendations about the
comparison of models, it doesn't establishes baselines or good practices
for training on imbalanced ecological data. Within the context of
networks, there are three specific issues that need to be adressed.
First, what values of performance measures are we expecting for a
classifier that has poor performance? This is particularly important as
it can evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training}, and can we
remedy it? Finally, because the low amount of data on interaction makes
a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
2015) hard to apply, which indicators can be optimized with the least
amount of positive interaction data?

@@

Chicco et al. (2021) - MCC maximizes other measures, other measures do
not maximize MCC, except notably when prevalence is low, or the baseline
guessing level is uncertain, which applies to networks. In this cases,
informedness should be used as a maximization criterion. Formulating
guidelines is particularly important, because most of the litterature
existing on optimizing classifier performance in the life sciences is
focused on genomics applications (which has very specific challenges,
see a recent discussion by Whalen et al., 2021), and can give
contradictory recommendations (Boughorbel et al., 2017; Chicco \&
Jurman, 2020; Delgado \& Tibau, 2019). This points not to a defficiency
in the litterature, but rather to a need for domain-specific evaluation
of how the particular ways in which datasets are biased can affect the
performance of predictive models.

@@

>>>>

\begin{itemize}
\tightlist
\item
  example on diagnostic test: rare events are hard to detect even with
  really good models
\item
  summary of model challenges for networks
\item
  Strydom et al. (2021) importance of drawing on traits + validation is
  challenging + comparing across space
\end{itemize}

@@

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence. In a previous
work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
of model performance than the True Skill Statistic (TSS), which we will
refer to as Youden's informedness (or \(J\)); these conclusions were
later criticized by Somodi et al. (2017), who emphasized that
informedness' relationship to prevalence depends on assumptions about
bias in the model, and therefore recommend the use of \(\kappa\) as a
validation of classification performance. Although this work offers
recommendations about the comparison of models, it doesn't establishes
baselines or good practices for training on imbalanced ecological data.
Within the context of networks, there are three specific issues that
need to be adressed. First, what values of performance measures are we
expecting for a classifier that has poor performance? This is
particularly important as it can evaluate whether low prevalence can
lull us into a false sense of predictive accuracy. Second, independently
of the question of model evaluation, is low prevalence an issue for
\emph{training}, and can we remedy it? Finally, because the low amount
of data on interaction makes a lot of imbalance correction methods (see
\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
optimized with the least amount of positive interaction data?

In addition to the literature on SDMs, most of the research on machine
learning application to life sciences is focused on genomics (which has
very specific challenges, see a recent discussion by Whalen et al.,
2021); this sub-field has generated largely different recommendations.
Chicco \& Jurman (2020) suggest using Matthews correlation coefficient
(MCC) over \(F_1\), as a protection against over-inflation of predicted
results; Delgado \& Tibau (2019) advocate against the use of Cohen's
\(\kappa\), again in favor of MCC, as the relative nature of \(\kappa\)
means that a worse classifier can be picked over a better one;
similarly, Boughorbel et al. (2017) recommend MCC over other measures of
performance for imbalanced data, as it has more desirable statistical
properties. More recently, Chicco et al. (2021) temper the apparent
supremacy of the MCC, by suggesting it should be replaced by Youden's
informedness (also known as \(J\), bookmaker's accuracy, and the
True-Skill Statistic) when the imbalance in the dataset may not be
representative (Jordano, 2016a, which is the case as networks are
under-sampled; 2016b; McLeod et al., 2021), when classifiers need to be
compared across different datasets, and when comparing the results to a
no-skill (baseline) classifier is important. As these conditions are
likely to be met with network data, there is a need to evaluate which
measures of classification accuracy respond in a desirable way.

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, amd can therefore return
somethins considered to be a pseudo-probability); in this case, the
optimal value below which predicitions are assumed to be negative
(\emph{i.e.} the interaction does not exist) can be determined by
picking a threshold maximizing some value on the ROC curve or the PR
curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
(2015) established that the ROC-AUC is biased towards over-estimating
performance for imbalanced data.

@@

@@

\leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%
Jordano, P. (2016a). Chasing Ecological Interactions. \emph{PLOS Biol},
\emph{14}(9), e1002559.
\url{https://doi.org/10.1371/journal.pbio.1002559}

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016b). Sampling networks of ecological interactions.
\emph{Functional Ecology}. \url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\hypertarget{ref-Saito2015PrePlo}{}%
Saito, T., \& Rehmsmeier, M. (2015). The Precision-Recall Plot Is More
Informative than the ROC Plot When Evaluating Binary Classifiers on
Imbalanced Datasets. \emph{PLOS ONE}, \emph{10}(3), e0118432.
\url{https://doi.org/10.1371/journal.pone.0118432}
